---
title: "Homework 24"
author: "Trent Meyer"
date: "2022-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This homework has us reading and working through the first half of Chapter 5 of the Getting Started with R book. We are working through to section 5.3.4, which looks at t-tests.

It said in the book to always start with visualizations, so you can have an expectation of what you want to see, and then you can do the statistical test.
1. plot your data 
2. build your model
3. check your assumptions
4. interpret your model
5. replot the data and the model

# 5.2 Chi-Squared Contingency Table Analysis

Looks at count data, and "is essentially a test of association among two or more categorical variables."

```{r}
# My First Chi-Square contingency analysis

# clear the environment
rm(list = ls()) 

# libraries I will use 
library(dplyr)
library(ggplot2)
library(here)

# import the data 
lady <- read.csv(here("Data", "ladybirds.csv"))

# check it out
glimpse(lady)
```

We want to create a table of the totals for each colour and location.
```{r}
totals <- lady %>%
  group_by(Habitat, colour) %>%
  summarise(total.number = sum(number))
```

```{r}
ggplot(totals, aes(x = Habitat, y = total.number,
                   fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c(black = "black", red = "red"))
```

fill changes the fill color of the bar, while colour changes the outline

stat = 'identity' tells ggplot to use just the data we are giving it. position = 'dodge' puts the two bars next to each other

```{r}
lady.mat <- xtabs(number ~ Habitat + colour,
                  data = lady)
lady.mat # let's us cross tabulate the data in a matrix
```


```{r}
lady.chi <- chisq.test(lady.mat)
lady.chi$expected
```

Very small p-value, (the probability we see is consistent with the null hypothesis), meaning that there is no association between the variables, and we choose to reject the null.

# 5.3.4 Two-sample t-test

Comparison of the means fo two groups of numeric values. It is appropriate for small sample sizes. Assumptions of constant variance and normality.

```{r}
ozone <- read.csv(here("Data", "ozone.csv"))
glimpse(ozone)
```

Plot the data to see expected relationship
```{r}
ggplot(ozone, aes(x = Ozone)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~ Garden.location, ncol = 1) +
  theme_bw()
```

We can see that the peaks are at different locations on the x-axis.

```{r}
t.test(Ozone ~ Garden.location, data = ozone)
```

We can see in the output that we are given a 95% CI, and it gives us the Ha. The p-value is small so we can reject the null hypothesis.

```{r}
var.test(Ozone ~ Garden.location, data = ozone)
```

# Conclusion

We learned how to use perform a chisquare and t-test. I think I will just need to get used to this whole process again, as it has been a while since I did this last.