---
title: "Homework23.Rmd"
author: "Trent Meyer"
date: "2022-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This homework is giving us practice with data visualizations, and the beginning of data analysis. We are creating a scatterplot, histogram, and box-and-whisker plot for two variables.

## Usual Code
```{r}
rm(list = ls()) # clears the environment
library(ggplot2) # loads in the ggplot2 library
library(here) # loads in the here library
library(dplyr) # loads in the dplyr library
```

```{r}
soil_df <- read.csv(here("Data", "soil_data.csv"))

library(stringr)
soil_df <- soil_df %>% mutate(field = word(Field.Name, 1))
```

## Scatterplot for SOM_percent
```{r}
ggplot(data = soil_df, aes(x = SOM_percent, y = field)) +
  geom_point()
```

Here I was a little unsure with the Field.Name, if it would be better to separate() this variable into variables called: field, plot, and sample.

I ended up using word() to get the field into a new variable, and just use that instead

## Histogram for SOM_percent
```{r}
ggplot(data = soil_df, aes(x = SOM_percent)) +
  geom_histogram()
```

## Box-and-Whisker Plot for SOM_percent
```{r}
ggplot(data = soil_df, aes(x = SOM_percent)) +
  geom_boxplot()
```

## Scatterplot for Ca.ppm
```{r}
ggplot(data = soil_df, aes(x = Ca.ppm, y = field)) +
  geom_point()
```

I ended up using word() to get the field into a new variable, and just use that instead

## Histogram for Ca.ppm 
```{r}
ggplot(data = soil_df, aes(x = Ca.ppm)) +
  geom_histogram()
```

## Box-and-Whisker Plot for Ca.ppm
```{r}
ggplot(data = soil_df, aes(x = Ca.ppm)) +
  geom_boxplot()
```

# Conclusion

I learned that it is possible to create a different plot compared to another person, and have both methods be correct. For the Field.Name, I chose to use word(), while someone else could have used separate, or they could have faceted.