---
title: "Homework-21.Rmd"
author: "Trent Meyer"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This homework is looking at GSWR Chapter 4: "Visualizing Your Data" and helps us look at how to make scatterplots, histograms, and box-and-whisker plots using dplyr and ggplot2.

```{r}
# plotting basics with ggplot

# libraries I need
library(dplyr) # loads in the dplyr package
library(ggplot2) # loads in the ggplot2 package
library(here) # loads in the here package

rm(list = ls()) # clears the environment

# read in the compensation data set
compensation <- read.csv(here("Data", "compensation.csv"))

# check out the data
glimpse(compensation)
```

# Plain and Simple bivariate scatterplot using ggplot2
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point()
```

First we give the data set as the first argument, and then we specify the aesthetics.

`+` symbol lets us add layers after the first layer, and we need to specify a geom, here we did geom_point() which added the points onto the first layer we made.

# Scatterplots 

## Getting rid of the grey background
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point() +
  theme_bw() # here is what changed
```

## Increase the size of the points
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point(size = 5) + # here is what changed
  theme_bw()
```

## Add axis labels
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point(size = 5) +
  xlab("Root Biomass") + # here is what changed
  ylab("Fruit Production") + # here is what changed
  theme_bw()
```

## Adjust colours to match levels
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, colour = Grazing)) + # here is what changed
  geom_point(size = 5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw()
```

## Change the shape of points
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, shape = Grazing)) + # here is what changed
  geom_point(size = 5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw()
```

# Box-and-whisker plots

Good for showing variation along with either the median or mean.

```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() +
  xlab("Grazing treatment") +
  ylab("Fruit Production") +
  theme_bw()
```

## Adding the raw data
```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() +
  geom_point(size = 4, colour = 'lightgray', alpha = 0.5) + # alpha blurs out the points
  xlab("Grazing treatment") +
  ylab("Fruit Production") +
  theme_bw()
```

Plants with wider root diameters produce more fruit compared to smaller diameters. Grazed plants produce more fruit compared to ungrazed.

# Histograms

```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram()
```

## Changing bin width
```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(bins = 10)
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(binwidth = 15)
```

# Faceting

Produces a matrix of graphs by a categorical variable in our data!

```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(binwidth = 15) + 
  facet_wrap(~Grazing)
```

# ggsave()

Let's us save a plot as a .png

```{r}
ggsave("ThatCoolHistogramofFruit.png")
```


# Conclusion

We learned how to make scatterplots, histograms, and box-and-whisker plots. We learned that the arguments are first data, and then we specify the aesthetics. We learned how to change axis labels, point shapes and sizes, colour of points, and how to add layers. We also learned about faceting, which is super helpful with categorical data.