---
title: "Meyer_Homework12"
author: "Trent Meyer"
date: "9/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework is practicing using dplyr functions and concepts from GSWR Chapters 2 and 3

```{r}
rm(list = ls()) # clear's the environment
library(dplyr) # loads in the dplyr package
library(here) # loads in the here package
Cover <- read.csv(here("Data", "donnerville_percent_cover.csv"), stringsAsFactors = TRUE) # reads in the percent cover df and call it Cover with strings as factors
```

```{r}
str(Cover) # look at the structure of Cover
head(Cover) # look at the first six observations in Cover
```

```{r}
Cover %>% filter(Site == "Donnerville 1") %>% group_by(Type) %>% summarise(avg_percent_cover = mean(Percent))
```

This is taking the Cover df, filter out for rows where Donnerville 1 is the site, grouping by the type of cover, and calculating the mean percent cover for the groups.

```{r}
Cover %>% filter(Type == "Vascular") %>% summarise(avg_percent_cover = mean(Percent))
```

This is taking the Cover df, filtering out for the Vascular type, and then determining the average percent vascular plant cover across all sites.