---
title: "Meyer_HW11_GSWRCh3Ex1"
author: "Trent Meyer"
date: "9/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction: Here I am working on Homework 11, which goes through Chapter 3 called "Data Management, Manipulation, and Exploration with `dplyr`" in the GSWR book.

# Setting Things Up
```{r}
rm(list = ls()) # clears the environment
library(ggplot2) # loads in the ggplot2 package
library(dplyr) # loads in the dplyr package
library(here) # loads in the here package
compensation <- read.csv(here("Data", "compensation.csv")) # loads in the compensation data set
```

# 3.1 Summary Statistics for each variable
```{r}
glimpse(compensation) # checking to make sure we have the correct data
summary(compensation) # get summary statistics for the compensation variables
```

The summary() function gives us a good idea if our data is extreme or doesn't make sense. A good thing to look at after reading your data into R.

# 3.2 dplyr verbs
`select()`, `slice()`, `filter()`, `arrange()`, and `mutate()` are five functions used to manipulate data from the dplyr package.

`select()`: selecting columns 
`slice()`: selecting rows
`filter()`: getting a subset of rows given a specified condition
`arrange()`: sorting rows
`mutate()`: creating new variables

First argument for all dplyr functions is the data frame!

# 3.3 Subsetting

`select()`
```{r}
select(compensation, Fruit) # returns the Fruit column in a data frame
```

select() asks for a data frame and then returns a data frame

```{r}
select(compensation, -Root) # returns everything except for Root in a data frame
```

`slice()` "works by returning specific row numbers you ask for"
```{r}
slice(compensation, 2) # returns the second row in compensation
```

returning sequences of rows
```{r}
slice(compensation, 2:10) # returns the second to the tenth rows
```

returning discontinuous sets of rows
```{r}
slice(compensation, c(2, 3, 10))
```

Also returns a data frame, however does not return the row number identity found in the original!!

```{r}
with(compensation, Fruit > 80) # produces a sequence of TRUE/FALSE saying where in Fruit the condition is met
```

`filter()`
```{r}
# find the rows where it is True that Fruit is >80 and return them as a df
filter(compensation, Fruit > 80)
```

`==` equals
`!=` does not equal
`>` greater than
`>=` greater than or equal to
`<` less than
`<=` less than or equal to 
`|` or
`&` and

```{r}
filter(compensation, Fruit > 80 | Fruit < 20) 
# returns rows where Fruit is greater than 80 OR less than 20
```

```{r}
lo_hi_fruit <- filter(compensation, Fruit > 80 | Fruit < 20)
lo_hi_fruit
```

# 3.4 Transforming

`mutate()`
```{r}
# what does compensation look like now
head(compensation)

# use mutate
# log(Fruit) is in the column logFruit
# all of which gets put into the object compensation
compensation <- mutate(compensation, logFruit = log(Fruit))

# first 6 rows of the new compensation
head(compensation)
```

By renaming a new data frame the same name as what it was originally called, you can overwrite it

# 3.5 Sorting

`arrange()`
```{r}
arrange(compensation, Fruit) # if we want to see the compensation df in increasing Fruit production order
```

You can use more than one dplyr function in one line of code
```{r}
select(filter(compensation, Fruit > 80), Root) # filter by Fruit production and the Root widths only
```

Reading from the inside out will help

# Piping

Makes things make more sense rather than putting many functions inside of each other

```{r}
compensation %>% 
  filter(Fruit > 80) %>% 
  select(Root)
```

Nested approach
```{r}
mean.fruit <- summarise(
  group_by(compensation, Grazing), # grouping by grazing
  meanFruit = mean(Fruit)) # calculate the mean for each of the grazing groups
```

Pipe approach
```{r}
compensation %>% 
  group_by(Grazing) %>%
  summarise(meanFruit = mean(Fruit),
            sdFruit = sd(Fruit))
```

Piping saves a lot of confusion, and lets you use the dplyr functions to easily manipulate data frames to do what you want!